### 2018-12-14 report/paper.tex
report

% \documentclass[sigplan,10pt,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
% \documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[sigplan,10pt]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmConference[Purdue CS590-DLS]{}{Dec 2019}{West Lafayette, Indiana}
\acmYear{2019}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.8]{beramono}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xcolor,colortbl}
\usepackage{url}
\usepackage{listings}
\usepackage{paralist}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{flushend}
\usepackage{bcprules}
\usepackage{nicefrac}
\usepackage{xfrac}
\input{macros}

\begin{document}

%% Title information
\title{Learning Heuristics of SAT by Graph Neural Networks}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
% \subtitle{Subtitle}                     %% \subtitle is optional
% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

\author
[Guannan Wei]
{Guannan Wei \\ Purdue University}

\lstMakeShortInline[keywordstyle=,%
              flexiblecolumns=false,%
              %basewidth={0.56em, 0.52em},%
              mathescape=false,%
              basicstyle=\tt]@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  In this document, we summarize the discussion and provide some formal description of the proposed model.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\renewcommand\thefootnotecopyrightpermission{}
\footnotetextcopyrightpermission{Copyright held by the authors.}
\fancyhead[RO,LE]{Preprint, Dec 2019}

\section{Problem Statement}

In this paper, we report the progress and result on the project that aims to learn branching heuristics 
for SAT problem based on DPLL algorithm. The DPLL algorithm \cite{Davis:1962:MPT:368273.368557} and its 
variants are widely used in many modern SAT solvers. 
The core ideas behind DPLL are searching and unit propagation. 
The searching involves choosing a variable and assigning it to either $\top$ or $\bot$, 
which is the part of our concern in this paper.

The DPLL algorithm works on Conjunction Normal Form. A CNF formula is a conjunction of clauses. A clause is a disjunction of literals. A literal is either a variable (symbol) or a negation of a variable. Formulas in other forms can be translated to CNF \cite{tseitin1983complexity}. For example, $(x_1 \vee x_2) \wedge (\neg x_2 \vee x_3)$ is a valid CNF formula.

The pseudocode of DPLL procedure is listed as follows:

\begin{lstlisting}
def dpll(f: Formula): Boolean = {
  if (f.hasUnsatClause) return false
  if (f.isSat) return true
  if (f.hasUnitClause) return dpll(f.elimUnitClause)
  val v = f.pickVariable
  val tryTrue = dpll(f.addUnitClause(v))
  if (tryTrue) true
  else dpll(f.addUnitClause(neg(v)))
}
\end{lstlisting}

Particularly, we would like to learn the branching heuristics for \texttt{f.pickVariable}.
In this paper, we use graph neural network to encode a CNF formula and use reinforcement learning 
to learn the best branching heuristics.
We use the policy gradient method for the reinforcement learning part.

\section{Proposed Model}

\subsection{General Description}

\paragraph{Graph Representation}

The graph embedding captures the permutation invariance and the structure of a CNF formula.
We first review NeuroSAT's graph embedding of a CNF formula \cite{selsam2018learning}.

Given a formula $F$ of $m$ clauses and $n$ variables in CNF, 
one can construct a graph $G = \langle V, E \rangle$ such that 
$V = V_{lit} \cup V_{cls}$ where $V_{lit}$ is the set of vertices embedding literals,
$V_{cls}$ is the set of vertices embedding clauses. 
And $E$ contains two different types of edge, the first one is the edegs connecting literals and its
containing clauses, the second one is the edges connecting a variable and its negation, formally,
$ E = \{ (l, c) | l \in c \} \cup \{ (l, \neg l) | l \}$ for all literals $l$ and clauses $c$.
Note that in the NeuroSAT's encoding, there is a vertex for every variable and its negation no matter
it appears or not, so the size of vertices set $|V| = 2n + m$.

\paragraph{Message-Passing Neural Network}

After the graph is constructed, NeuroSAT uses message-passing mechanism to refine the vector space embeddings.
One iteration of a message-passing consists of two stages:
1) first every clause receives messages from its neighboring literals and update the clause embedding;
2) then every literal receives its message from its neighboring clauses and its negated literal vertices,
and update its embedding.

Suppose the size of embedding vectors is $d$. The initial embedding vectors of literals and clauses 
are $L^{0} \in \mathbb{R}^{2n \times d}$ and $C^{0} \in \mathbb{R}^{m \times d}$ which are randomly initialized.
NeuroSAT uses two MLPs $\mathcal{L}_{msg}$ and $\mathcal{C}_{msg}$ to produce the messages from embeddings, 
and two LSTMs $\mathcal{L}_u$ and $\mathcal{C}_u$ to update the embeddings every iteration 
(the hidden states are elided). 
As mentioned above, for the time $t$ there are two stages in a iteration:
$$ C^{t+1} \leftarrow \mathcal{C}_u( \mathcal{L}_{msg}(L^t) ) $$
$$ L^{t+1} \leftarrow \mathcal{L}_u( \mathcal{C}_{msg}(C^{t+1}) ) $$

After $T$ iterations, with the embedding produced, another MLP $\mathcal{P}$ which served as the policy network takes 
the embeddings as input and produces the vector $\mathcal{P}(L^T) \in \mathbb{R}^{2n}$ that represents the probabilities of choosing which literal.

\subsection{Embedding Problem}

However, naively adapting NeuroSAT's graph embedding is not going to work.
Because NeuroSAT models SAT as a supervised learning problem without using traditional algorithms 
such as DPLL, which then means the graph structure is static. 
But in the DPLL and reinforcement learning setting, the formula is changed every time when we make a decision --
some variables and clauses may completely be eliminated by unit propagation over the time.

This further raises two entangled problems:
1) How to embed the assignments as well as the decision history into the graph.
2) How to reflect the structural changes in the graph representation. Ideally, if the assignments are properly 
embedded, the graph should be able to automatically reflect the resulting formula after an assignment. 
Otherwise, adding annotations on nodes to identify live literals and clauses or manually pruning the graph 
(i.e., removing the edges and nodes) are two options.

\paragraph{Solutions}

It seems a perfect self-adapting graph neural network architecture that reflects the assignments
does not exist yet.

So we discuss the second pruning schema that changes the graph structure every time.
Note that there are conceptually two neural networks, $\mathcal{G}$ transforms a formula (graph represented) 
to embeddings, and $\mathcal{P}$ predicts how to choose from the live variables according to the embeddings.
Also note that if conflict-driven clauses learning is not introduced, 
the graph is monotonically shrinking every time.

Consider a complete decision trace that consists of $t$ decisions:
$$ (F_0, G_0, d_0) \to (F_1, G_1, d_1) \cdots \to (F_t, G_t, d_t) $$

where $F_i$ is the formula, $G_i$ is the corresponding graph embedding, and $d_i$ is the decision at time $i$.
A decision $d$ is a pair that consists of a variable and a assigned value (either $\top$ or $\bot$).

If eventually we have SAT, then we assign a reward $1$ for every decision on the this path,
and backprop the two neural networks $\mathcal{G}$ and $\mathcal{P}$ at each level of decisions.

If eventually we have UNSAT or have to backtrack, then we assign reward $-\frac{1}{2^{t-i+1}}$
and backprop the $\mathcal{G}$ and $\mathcal{P}$ at the level $i$.
For example, we assign the most recent decision a 
reward $-\frac{1}{2}$ and backprop the $\mathcal{G}$ and $\mathcal{P}$ at this level; then assign
a reward $-\frac{1}{4}$ and backprop the $\mathcal{G}$ and $\mathcal{P}$ at level $t-1$, and etc.

\clearpage{}

% %% Acknowledgments
% \begin{acks}                            %% acks environment is optional
%                                         %% contents suppressed with 'anonymous'
%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%   %% acknowledge financial support and will be used by metadata
%   %% extraction tools.
%   This material is based upon work supported by the
%   \grantsponsor{GS100000001}{National Science
%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%   conclusions or recommendations expressed in this material are those
%   of the author and do not necessarily reflect the views of the
%   National Science Foundation.
% \end{acks}

%% Bibliography
\bibliography{reference}

%% Appendix
% \appendix
% \section{Appendix}

% Text of appendix \ldots

\end{document}

### 2018-12-14 report/paper.tex
update

@@
 %% before \maketitle command
-\begin{abstract}
-  In this document, we summarize the discussion and provide some formal description of the proposed model.
-\end{abstract}
+%\begin{abstract}
+%\end{abstract}
 
@@
 
-\section{Problem Statement}
+\section{Introduction}
 
-In this paper, we report the progress and result on the project that aims to learn branching heuristics 
-for SAT problem based on DPLL algorithm. The DPLL algorithm \cite{Davis:1962:MPT:368273.368557} and its 
-variants are widely used in many modern SAT solvers. 
-The core ideas behind DPLL are searching and unit propagation. 
-The searching involves choosing a variable and assigning it to either $\top$ or $\bot$, 
-which is the part of our concern in this paper.
+In this paper, we report the progress and preliminary result on the CS590-DLS course project that aims 
+to learn branching heuristics for the Boolean Satisfiability (SAT) problem based on DPLL algorithm.
 
-The DPLL algorithm works on Conjunction Normal Form.
-A CNF formula is a conjunction of clauses.
-A clause is a disjunction of literals.
+\paragraph{Background}
+
+Given a formula consists of variables over booleans and propositional logic connectives, the satisfiability
+problem asks an assignment, i.e., a mapping from variables to boolean values, that makes the formula be true. 
+For example; $(P \wedge Q) \to (\neg Q \vee R)$ is a well-formed formula and can be satisfied by assigning
+$P$ to be $\top$; $P \wedge \neg P$ is another well-formed formula but can not be satisfied, written as UNSAT.
+
+A traditional approach to solve SAT is DPLL algorithm \cite{Davis:1962:MPT:368273.368557}. 
+Although in the worst case, DPLL is still an exponential algorithm due the nature of the problem,
+but on many instances DPLL is able to find a solution or prove UNSAT quickly.
+Thus DPLL and its variants are widely used in many modern SAT solvers such as miniSAT.
+
+The DPLL algorithm works on Conjunction Normal Form (CNF). 
+A CNF formula is a conjunction of clauses. A clause is a disjunction of literals. 
 A literal is either a variable (symbol) or a negation of a variable. 
 Formulas in other forms can be translated to CNF \cite{tseitin1983complexity}. 
-For example, $(x_1 \vee x_2) \wedge (\neg x_2 \vee x_3)$ is a valid CNF formula.
+For example, $(x_1 \vee x_2) \wedge (\neg x_2 \vee x_3)$ is a well-formed formula in CNF.
+
+With the CNF format introduced, now we can briefly describe the essential ideas behind DPLL: 
+unit propagation, searching and backtracking.
+If a clause contains only one literal, then it is a unit clause. Since the problem is given 
+in CNF, this cluase is forced to be satisfied, thus we must assign this literal to be true. 
+The unit propagation then calculate the consequential formula by making this assignment --
+some clauses can be complete eliminated.
+If the current formula does not contain an unit clause, the algorithm then chooses a branching
+variable by predefined heuristics and assign either $\top$ or $\bot$ to this variable,
+and then propagate this decision and recursively call the decision procedure.
+At some point, we may reach a dead end that the current assignment shows an UNSAT result, 
+then we need to recover to some previous decision we made and try another assignment.
 
@@
 
+Several variants based on this algorithm make it more efficient: 1) one can introduce
+pure literal elimination, that is a variable appears with only one polarity can be 
+assigned to a value that makes its containing clauses satisfies but without triggering
+any conflict. 2) When we reach one dead end, we can perform a conflict analysis which 
+produces a lemma that must be satisfied (known as conflict-driven clauses learning, CDCL
+\cite{DBLP:series/faia/SilvaLM09}). 
+The lemma will be added back to the formula, and the backtrack can be done in a non-chronological fashion.
+In this paper, we constrain ourself to the simplest version of DPLL and focus on
+learning how to choose the branching variable.
+
+\paragraph{Problem Statement}
+
+The fundamental problem we are interested in is
+
 Particularly, we would like to learn the branching heuristics for \texttt{f.pickVariable}.
-In this paper, we use graph neural network to encode a CNF formula and use reinforcement learning 
-to learn the best branching heuristics.
+In this paper, we propose to use recent advances in graph neural network to 
+encode a CNF formula and use reinforcement learning to learn the best branching heuristics.
 We use the policy gradient method for the reinforcement learning part.
 
-\section{Proposed Model}
+\section{Model}
+
+In this section, we briefly review the related work on graph representations of SAT
+and then describe the model used in this paper.
 
@@
 
+\section{Implementation}
+
+To make the DPLL algorithm integrates well with the ML/RL part, we implement a small-step version of 
+DPLL, but without changing the decision procedure. This can be obtained by applying 
+defunctionalization \cite{DBLP:conf/ppdp/DanvyN01} to the recursive DPLL implementation, which exposed a continuation
+component that saves the decision history and the current state. Another similar presentation
+is \citeauthor{DBLP:conf/lpar/NieuwenhuisOT04}'s abstract DPLL.
+This allows us easily inspect the current state and explicitly handle backtrack.
+
+\section{Experiment}
+
+\section{Related Work}
+
+As artifical neural networks greatly succeeded in vision and natural language tasks in last ten years,
+many works that combine deep learning and logical reasoning also has been emerged recently.
+
+\section{Conclusion}
+
+\textbf{Acknowledgement} The author would like to thank Fei Wang for the inspiring discussion and 
+help of the machine learning part.
+
 \clearpage{}

### 2018-12-14 rein/mpnn.py
mpnn

@@ -1,0 +1,204 @@
+import os
+import time
+import argparse
+import numpy as np
+
+import torch
+import torch.nn as nn
+from torch.autograd import Variable
+import torch.nn.functional as F
+
+#TODO: batch for a NNet
+
+dtype = torch.FloatTensor
+
+# Reference: https://github.com/priba/nmp_qc/
+
+class NNet(nn.Module):
+    def __init__(self, n_in, n_out, hlayers = (128, 258, 128)):
+        super(NNet, self).__init__()
+        self.n_hlayers = len(hlayers)
+        self.fcs = nn.ModuleList([nn.Linear(n_in, hlayers[i]) if i == 0 else
+                                  nn.Linear(hlayers[i-1], n_out) if i == self.n_hlayers else
+                                  nn.Linear(hlayers[i-1], hlayers[i])
+                                  for i in range(self.n_hlayers+1)])
+    def forward(self, x):
+        x = x.contiguous().view(-1, self.num_flat_features(x))
+        for i in range(self.n_hlayers):
+            x = F.relu(self.fcs[i](x))
+        x = self.fcs[-1](x)
+        return x
+    def num_flat_features(self, x):
+        size = x.size()[1:]
+        num_features = 1
+        for s in size: num_features *= s
+        return num_features
+
+class MessageFunction(nn.Module):
+    def __init__(self, message_def='mpnn', args={}):
+        super(MessageFunction, self).__init__()
+        self.m_definition = ''
+        self.m_function = None
+        self.args = {}
+        self.__set_message(message_def, args)
+    def forward(self, h_v, h_w, e_vw, args=None):
+        return self.m_function(h_v, h_w, e_vw, args)
+    def __set_message(self, message_def, args={}):
+        self.m_definition = message_def.lower()
+        self.m_function = {'mpnn': self.m_mpnn }.get(self.m_definition, None)
+        if self.m_function is None:
+            raise RuntimeError("Can not find message function for {}".format(self.m_definition))
+        init_parameters = { 'mpnn': self.init_mpnn }.get(self.m_definition, lambda x: (nn.ParameterList([]),
+                                                                                       nn.ModuleList([]),
+                                                                                       {}))
+        self.learn_args, self.learn_modules, self.args = init_parameters(args)
+        self.m_size = { 'mpnn': self.out_mpnn }.get(self.m_definition, None)
[…]

### 2018-12-15 rein/reinforce.py
misc

@@ -2,7 +2,10 @@ import argparse
 import gym
 import numpy as np
 from itertools import count
-
+import os
+import random
+import itertools
+import bisect
 import torch
 import torch.nn as nn
 import torch.nn.functional as F
@@ -25,7 +28,8 @@ args = parser.parse_args()
 def apply_backtrack(s: State) -> State:
     #print("backtrack {}".format(s))
     v, f, asn = s.cont[0]
-    return State(f.assign(v, False), (-v,)+asn, s.cont[1:])
+    b = False if v>0 else True
+    return State(f.assign(v, b), (-v,)+asn, s.cont[1:])
 
 def apply_unit(s: State) -> State:
     #print("unit {}".format(s))
@@ -50,7 +54,7 @@ class SatEnv():
         if len(cs) < 91: return torch.tensor(cs + ([([0.0]*3)] * (91-len(cs)))).flatten()
         else: return torch.tensor(cs).flatten()
     def action_space(self) -> List[Lit]:
-        return [abs(x) for x in self.state.formula.allVars]
+        return self.state.formula.allVars
     def step_tr(self):
         f, asn, cont = self.state
         if f.isEmpty(): return True
@@ -74,33 +78,38 @@ class SatEnv():
         f, asn, cont = self.state
         if f.isEmpty(): 
             reward = 0.0
+            print("pick {}, done: {}, backtrack/unsat: {}".format(action, True, False))
             return None, reward, True, asn
         elif len(cont) == 0 and f.hasUnsat():
-            reward = 0.0
+            reward = -1.0
+            print("pick {}, done: {}, backtrack/unsat: {}".format(action, True, True))
             return None, reward, True, None
         elif f.hasUnsat():
             self.state = apply_backtrack(self.state)
-            reward = 0.0
+            reward = -1.0
             done = True if len(self.state.formula.cs) == 0 else False
+            print("pick {}, done: {}, backtrack/unsat: {}".format(action, done, True))
             return self.emb(self.state.formula), reward, done, self.state.assignment
         elif f.hasUnit():
             self.state = apply_unit(self.state)
             return self.step(action)
-        v = action
+        v = abs(action)
+        b = True if action>0 else False
         pre_nvars = len(self.state.formula.allVars)
-        self.state = State(f.assign(v, True), (v,)+asn, (Cont(v, f, asn),)+cont)
+        self.state = State(f.assign(v, b), (action,)+asn, (Cont(action, f, asn),)+cont)
         cur_nvars = len(self.state.formula.allVars)
[…]

### 2018-12-15 report/paper.tex
misc

@@
 
-The fundamental problem we are interested in is
+In the literatures, people have invented several efficient branching heuristics, 
+from very simple ones such as picking the most frequent literal, 
+to complicated ones such as such as VSIDS \cite{zhang2002quest}.
+Literatures on branching heuristics can be found from \cite{hooker1995branching, li1997heuristics}.
+But among these heuristics, no one works perfect for all kinds of instances.
 
-Particularly, we would like to learn the branching heuristics for \texttt{f.pickVariable}.
-In this paper, we propose to use recent advances in graph neural network to 
-encode a CNF formula and use reinforcement learning to learn the best branching heuristics.
-We use the policy gradient method for the reinforcement learning part.
+The fundamental problem we are interested in is, given a such simple DPLL algorithm
+framework, can we learn an efficient heuristics from scratch?
+In this paper, we propose to use recent advances in graph neural networks to 
+encode a CNF formula and use reinforcement learning to learn the branching heuristics.
+We will use the policy gradient method for the reinforcement learning part.
 
@@
 
-The graph embedding captures the permutation invariance and the structure of a CNF formula.
+The central problem on apply neural-network-based approach to SAT is how to 
+represent a boolean formula and a decision in a differentiable way.
+The formula can be easily represented by a sequence or a matrix, but this
+representation losts the structural information and the invarant properties such as
+exchange.
+The graph neural networks have been applied to computational chemistry and 
+social network analysis which also capture the permutation invariance and the 
+structure of a CNF formula.
 We first review NeuroSAT's graph embedding of a CNF formula \cite{selsam2018learning}.
@@
 one can construct a graph $G = \langle V, E \rangle$ such that 
-$V = V_{lit} \cup V_{cls}$ where $V_{lit}$ is the set of vertices embedding literals,
-$V_{cls}$ is the set of vertices embedding clauses. 
-And $E$ contains two different types of edge, the first one is the edegs connecting literals and its
-containing clauses, the second one is the edges connecting a variable and its negation, formally,
+$V = V_{lit} \cup V_{cls}$ where $V_{lit}$ is the set of vertices that embed literals,
+$V_{cls}$ is the set of vertices that embed clauses. 
+And $E$ contains two different types of edge, the first one is the edegs that connect literals and its
+containing clauses, the second type is the edges that connect a variable and its negation, formally,
 $ E = \{ (l, c) | l \in c \} \cup \{ (l, \neg l) | l \}$ for all literals $l$ and clauses $c$.
@@
 
-After the graph is constructed, NeuroSAT uses message-passing mechanism to refine the vector space embeddings.
+After the graph is constructed, NeuroSAT uses iterative message-passing mechanism to refine the 
+vector space embeddings.
 One iteration of a message-passing consists of two stages:
@@
 
-After $T$ iterations, with the embedding produced, another MLP $\mathcal{P}$ which served as the policy network takes 
-the embeddings as input and produces the vector $\mathcal{P}(L^T) \in \mathbb{R}^{2n}$ that represents the probabilities of choosing which literal.
+After $T$ iterations, with the embedding produced, another MLP $\mathcal{P}$ which served as the policy 
+network takes the embeddings as input and produces the vector $\mathcal{P}(L^T) \in \mathbb{R}^{2n}$ that 
+represents the probabilities of choosing which literal.
 
@@
 
-\paragraph{Solutions}
+\paragraph{Potential Solutions}
 
@@
 
-To make the DPLL algorithm integrates well with the ML/RL part, we implement a small-step version of 
-DPLL, but without changing the decision procedure. This can be obtained by applying 
+To make the DPLL algorithm interactively integrates with the reinforcement learning algorithm, we implement a small-step version of 
+DPLL (as opposed to the recursive one in we showed in Section 1), but without changing the decision procedure. This can be obtained by applying 
 defunctionalization \cite{DBLP:conf/ppdp/DanvyN01} to the recursive DPLL implementation, which exposed a continuation
 component that saves the decision history and the current state. Another similar presentation
-is \citeauthor{DBLP:conf/lpar/NieuwenhuisOT04}'s abstract DPLL.
+is \citeauthor{DBLP:conf/lpar/NieuwenhuisOT04}'s abstract DPLL formulation \cite{DBLP:conf/lpar/NieuwenhuisOT04}.
 This allows us easily inspect the current state and explicitly handle backtrack.
 
+We use PyTorch framework to implement the neural networks and reinforcement learning.
+The code we use to implement message-passing neural network and gated-graph neural network 
+are based on open-sourced repositories on Github 
+\footnote{https://github.com/JamesChuanggg/ggnn.pytorch} 
+\footnote{https://github.com/priba/nmp\_qc}.
+The policy network is a shallow neural network with three hidden layers.
+
 \section{Experiment}
+
+In the experiments, we use two baselines: the first one is randomly picking a literal; 
+the second one we use a multi-layer perceptrons as policy network, but just with 
+a linear, direct embedding of CNF formula.
+
+The training data and testing data are public and available online 
+\footnote{https://www.cs.ubc.ca/~hoos/SATLIB/benchm.html}.
+To measure the efficiency, we count the number of decisions made by baselines and the 
+graph-based model.
 
@@
 \section{Conclusion}
+
+We investigate the problem of learning branching heuristics of SAT by 
+graph neural networks and show the preliminary result. 
+We note that since the random 3-SAT problems are not tightly constrained,
+which means there may exist many satisfiable assignments, the numbers we showed
+in Section 4 does not strictly imply the policy network has learned effective
+branching heuristics. We leave the examination of ability of generalizing
+on larger instances as future works. The graph currently experimented is still 
+static, implementing and tweaking the dynamicaly updated GNN (as discussed in 
+Section 2.2) is another on-going work.
 

### 2018-12-15 rein/ggnn.py

@@ -83,7 +83,7 @@ class GGNN(nn.Module):
         for m in self.modules():
             if isinstance(m, nn.Linear):
                 m.weight.data.normal_(0.0, 0.02)
-                m.bias.daa.fill_(0)
+                m.bias.data.fill_(0)
 
     def forward(self, prop_state, annotation, A):
         for i_step in range(self.n_steps):

### 2018-12-15 rein/mpnn.py

@@ -8,8 +8,6 @@ import torch.nn as nn
 from torch.autograd import Variable
 import torch.nn.functional as F
 
-#TODO: batch for a NNet
-
 dtype = torch.FloatTensor
 
 # Reference: https://github.com/priba/nmp_qc/
@@ -202,3 +200,4 @@ if __name__ == '__main__':
     print(net)
     print(rin)
     print(net(rin))
+

### 2018-12-15 report/paper.tex
misc

@@
 The code we use to implement message-passing neural network and gated-graph neural network 
-are based on open-sourced repositories on Github 
+are adapted from open-sourced repositories on Github 
 \footnote{https://github.com/JamesChuanggg/ggnn.pytorch} 
-\footnote{https://github.com/priba/nmp\_qc}.
+\footnote{https://github.com/yujiali/ggnn}.
 The policy network is a shallow neural network with three hidden layers.

### 2018-12-15 rein/ggnn.py

@@ -50,9 +50,9 @@ class Propogator(nn.Module):
 class GGNN(nn.Module):
     def __init__(self, opt):
         super(GGNN, self).__init__()
-        assert(opt.state_dim >= opt.annotation_dim)
+        #assert(opt.state_dim >= opt.annotation_dim)
         self.state_dim = opt.state_dim
-        self.annotation_dim = opt.annotation_dim
+        #self.annotation_dim = opt.annotation_dim
         self.n_edge_types = opt.n_edge_types
         self.n_node = opt.n_node
         self.n_steps = opt.n_steps
@@ -72,7 +72,9 @@ class GGNN(nn.Module):
 
         # Output model
         self.out = nn.Sequential(
-            nn.Linear(self.state_dim + self.annotation_dim, self.state_dim),
+            #nn.Linear(self.state_dim + self.annotation_dim, self.state_dim),
+            nn.Linear(self.state_dim, self.state_dim),
+            nn.Linear(self.state_dim, self.state_dim),
             nn.Tanh(),
             nn.Linear(self.state_dim, 1)
         )
@@ -85,7 +87,8 @@ class GGNN(nn.Module):
                 m.weight.data.normal_(0.0, 0.02)
                 m.bias.data.fill_(0)
 
-    def forward(self, prop_state, annotation, A):
+    #def forward(self, prop_state, annotation, A):
+    def forward(self, prop_state, A):
         for i_step in range(self.n_steps):
             in_states = []
             out_states = []
@@ -93,11 +96,13 @@ class GGNN(nn.Module):
                 in_states.append(self.in_fcs[i](prop_state))
                 out_states.append(self.out_fcs[i](prop_state))
             in_states = torch.stack(in_states).transpose(0, 1).contiguous()
+            print(in_states.shape)
             in_states = in_states.view(-1, self.n_node*self.n_edge_types, self.state_dim)
             out_states = torch.stack(out_states).transpose(0, 1).contiguous()
             out_states = out_states.view(-1, self.n_node * self.n_edge_types, self.state_dim)
             prop_state = self.propogator(in_states, out_states, prop_state, A)
-        join_state = torch.cat((prop_state, annotation), 2)
-        output = self.out(join_state)
+        #join_state = torch.cat((prop_state, annotation), 2)
+        #output = self.out(join_state)
+        output = self.out(prop_state)
         output = output.sum(2)
         return output

### 2018-12-15 rein/reinforce.py

@@ -103,7 +103,6 @@ class SatEnv():
         print("pick {}, done: {}, backtrack/unsat: {}".format(action, done, False))
         return self.emb(self.state.formula), reward, done, self.state.assignment
 
-# TODO: replace it as a GNN
 class Policy(nn.Module):
     def __init__(self):
         super(Policy, self).__init__()

### 2018-12-15 rein/gnn.py
gnn

@@ -1,0 +1,182 @@
+import numpy as np
+import torch
+import torch.nn as nn
+import torch.nn.utils
+from torch.autograd import Variable
+import torch.nn.functional as F
+from torch.nn.utils.rnn import pad_packed_sequence, pack_padded_sequence
+
+from typing import List, Tuple, Dict, Sequence, Any
+
+
+class AdjacencyList:
+    """represent the topology of a graph"""
+    def __init__(self, node_num: int, adj_list: List, device: torch.device):
+        self.node_num = node_num
+        self.data = torch.tensor(adj_list, dtype=torch.long, device=device)
+        self.edge_num = len(adj_list)
+
+    @property
+    def device(self):
+        return self.data.device
+
+    def __getitem__(self, item):
+        return self.data[item]
+
+
+class GatedGraphNeuralNetwork(nn.Module):
+    def __init__(self, hidden_size, num_edge_types, layer_timesteps,
+                 residual_connections,
+                 state_to_message_dropout=0.3,
+                 rnn_dropout=0.3,
+                 use_bias_for_message_linear=True):
+
+        super(GatedGraphNeuralNetwork, self).__init__()
+
+        self.hidden_size = hidden_size
+        self.num_edge_types = num_edge_types
+        self.layer_timesteps = layer_timesteps
+        self.residual_connections = residual_connections
+        self.state_to_message_dropout = state_to_message_dropout
+        self.rnn_dropout = rnn_dropout
+        self.use_bias_for_message_linear = use_bias_for_message_linear
+
+        # Prepare linear transformations from node states to messages, for each layer and each edge type
+        # Prepare rnn cells for each layer
+        self.state_to_message_linears = []
+        self.rnn_cells = []
+        for layer_idx in range(len(self.layer_timesteps)):
+            state_to_msg_linears_cur_layer = []
+            # Initiate a linear transformation for each edge type
+            for edge_type_j in range(self.num_edge_types):
+                # TODO: glorot_init?
+                state_to_msg_linear_layer_i_type_j = nn.Linear(self.hidden_size, self.hidden_size, bias=use_bias_for_message_linear)
+                setattr(self,
+                        'state_to_message_linear_layer%d_type%d' % (layer_idx, edge_type_j),
+                        state_to_msg_linear_layer_i_type_j)
+
+                state_to_msg_linears_cur_layer.append(state_to_msg_linear_layer_i_type_j)
[…]

### 2018-12-15 rein/reinforce_graph.py

@@ -1,0 +1,308 @@
+import argparse
+import gym
+import numpy as np
+from itertools import count
+import os
+import random
+import itertools
+import bisect
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import torch.optim as optim
+from torch.distributions import Categorical
+import glob
+from common import *
+from gnn import *
+
+parser = argparse.ArgumentParser(description='PyTorch REINFORCE example')
+parser.add_argument('--gamma', type=float, default=0.99, metavar='G',
+                    help='discount factor (default: 0.99)')
+parser.add_argument('--task_id', type=int, default=4, help='bAbI task id')
+parser.add_argument('--question_id', type=int, default=0, help='question types')
+parser.add_argument('--workers', type=int, help='number of data loading workers', default=2)
+parser.add_argument('--batchSize', type=int, default=20, help='input batch size')
+parser.add_argument('--state_dim', type=int, default=40, help='GGNN hidden state size')
+parser.add_argument('--n_steps', type=int, default=5, help='propogation steps number of GGNN')
+parser.add_argument('--niter', type=int, default=5, help='number of epochs to train for')
+parser.add_argument('--lr', type=float, default=0.01, help='learning rate')
+parser.add_argument('--cuda', action='store_true', help='enables cuda')
+parser.add_argument('--verbal', action='store_true', help='print training info or not')
+parser.add_argument('--seed', type=int, help='manual seed', default=543, metavar='N')
+opt = parser.parse_args()
+
+
+def apply_backtrack(s: State) -> State:
+    #print("backtrack {}".format(s))
+    v, f, asn = s.cont[0]
+    b = False if v>0 else True
+    return State(f.assign(v, b), (-v,)+asn, s.cont[1:])
+
+def apply_unit(s: State) -> State:
+    #print("unit {}".format(s))
+    f, asn, cont = s
+    new_f, new_asn = f.elimUnit()
+    return State(new_f, new_asn + asn, cont)
+
+def create_adjacency_matrix(edges, n_nodes, n_edge_types):
+    a = np.zeros([n_nodes, n_nodes * n_edge_types * 2])
+    for edge in edges:
+        src_idx = edge[0]
+        e_type = edge[1]
+        tgt_idx = edge[2]
+        a[tgt_idx-1][(e_type - 1) * n_nodes + src_idx - 1] =  1
+        a[src_idx-1][(e_type - 1 + n_edge_types) * n_nodes + tgt_idx - 1] =  1
+    return a
+
+class SatEnv():
[…]

### 2018-12-15 rein/gnn.py
gnn

@@ -112,15 +112,16 @@ class GatedGraphNeuralNetwork(nn.Module):
             # Extract residual messages, if any:
             layer_residual_connections = self.residual_connections.get(layer_idx, [])
             # List[(V, D)]
-            layer_residual_states: List[torch.FloatTensor] = [node_states_per_layer[residual_layer_idx]
-                                                              for residual_layer_idx in layer_residual_connections]
+            layer_residual_states = [node_states_per_layer[residual_layer_idx] for residual_layer_idx in layer_residual_connections]
 
             # Record new states for this layer. Initialised to last state, but will be updated below:
             node_states_for_this_layer = node_states_per_layer[-1]
             # For each message propagation step
             for t in range(num_timesteps):
-                messages: List[torch.FloatTensor] = []  # list of tensors of messages of shape [E, D]
-                message_source_states: List[torch.FloatTensor] = []  # list of tensors of edge source states of shape [E, D]
+                #messages: List[torch.FloatTensor] = []  # list of tensors of messages of shape [E, D]
+                messages = []  # list of tensors of messages of shape [E, D]
+                #message_source_states: List[torch.FloatTensor] = []  # list of tensors of edge source states of shape [E, D]
+                message_source_states = []  # list of tensors of edge source states of shape [E, D]
 
                 # Collect incoming messages per edge type
                 for edge_type_idx, adjacency_list_for_edge_type in enumerate(adjacency_lists):
@@ -138,7 +139,7 @@ class GatedGraphNeuralNetwork(nn.Module):
                         message_source_states.append(edge_source_states)
 
                 # shape [M, D]
-                messages: torch.FloatTensor = torch.cat(messages, dim=0)
+                messages = torch.cat(messages, dim=0)
 
                 # Sum up messages that go to the same target node
                 # shape [V, D]

### 2018-12-15 rein/reinforce_graph.py

@@ -113,7 +113,7 @@ class SatEnv():
 
         output = gnn.compute_node_representations(initial_node_representation=torch.randn(n_nodes, 32),
                                                   adjacency_lists=[lit_edges, clause_edges])
-        print("output shape: {}".format(output.shape))
+        #print("output shape: {}".format(output.shape))
         return output[:len(all_var_uniq)*2]
 
     def action_space(self) -> List[Lit]:

### 2018-12-15 report/paper.tex
report data

@@
 \usepackage{xfrac}
+\usepackage{graphicx}
 \input{macros}
@@
 We use PyTorch framework to implement the neural networks and reinforcement learning.
+The model we implemented is a gated-graph neural network, which is a variant of message-passing neural network.
 The code we use to implement message-passing neural network and gated-graph neural network 
@@
 
+As the first step, we train and test our models with 3-SAT problems consist of 
+20 variables and 91 clauses. The training set contains 100 instances; we test
+the models on 50 instances. For each training instance, we run 10 episodes and 
+setup up a upper bound 400 steps for each episodes; but usually one episodes 
+will finish in less than 100 steps.
 The training data and testing data are public and available online 
 \footnote{https://www.cs.ubc.ca/~hoos/SATLIB/benchm.html}.
-To measure the efficiency, we count the number of decisions made by baselines and the 
-graph-based model.
+To measure the efficiency, we count the number of decisions made by the DPLL algorithms 
+with different baselines and the graph-based model. 
+The intermediate states of DPLL such as unit propagation do not count as decisions.
+
+Table 1 shows the basic statistical information (number of decisions) of 50 testing instances.
+As we can see, compared with randomly picking a variable, 
+the naive linear embedding already improves the performance.
+However, we observe decline when using the graph neural network -- it has roughly the same
+performance with the random stategy according the average stesp.
+
+Figure 1 provides a detailed comparision between the three models based on the residual 
+steps of these 50 test instances. In the graph noted in the form of "A vs. B", we obtain
+the residual steps by computing $steps(B) - steps(A)$ for each instances.
+For example, Figure 1 (a) shows the comparision between random strategy and RL with linear 
+embedding, the negative numbers on y-axis mean that on these instances the RL model performs
+better than the random strategy -- overall, RL is better on 52\% instances.
+If we compare the random baseline with graph-based RL, again, we do not observe
+significant improvement. We conjuecture this happens from two reasons: 
+1) the graph model although captures the structural information, but it is more 
+sophisticated, further parameter tunning (as well as for the policy network) 
+may improve the performance;
+2) If the model is trained with more data, we can expect a performance imporvement.
+
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|}
+\hline
+           & mean  & std   & median \\ \hline
+random strategy    & 24.66 & 22.56 & 21.5   \\ \hline
+linear embedding RL & 19.94 & 12.90 & 15.0  \\ \hline
+graph embedding RL  & 24.88 & 17.79 & 19.5  \\ \hline
+\end{tabular}
+\caption{Basic statistical information (steps)}
+\end{table}
+
+\begin{figure}[ht!]
+  \centering
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \includegraphics[scale=0.3]{random_vs_linear.png}
+    \caption{random strategy vs. linear embedding RL}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+  \includegraphics[scale=0.3]{random_vs_graph.png}
+  \caption{random strategy vs. graph embedding RL}
+  \end{subfigure}
+
+  \caption{Detailed comparison of three models on 50 test instances.}
+\end{figure}
 
@@
 many works that combine deep learning and logical reasoning also has been emerged recently.
+Some of the work are completely based on neural network, as opposed to utilizing
+an existing deterministic framework. \citeauthor{selsam2018learning} proposed to 
+use message-passing neural network to learn the assignment of a SAT problem \cite{selsam2018learning}.
+We adapt a similar graph representation from \citeauthor{selsam2018learning}, but use it 
+on DPLL, which poses a difficulty on handling dynamic graph strcuture.
+Based on \citeauthor{selsam2018learning}'s NeuraSAT, a ICLR 2019 paper
+further improves the model that directly generates the assignment in a unsupervised fashion, 
+but works on circuit-sat problem\cite{anonymous2019learning}.
 

### 2018-12-15 report/paper.tex
update

@@
 However, we observe decline when using the graph neural network -- it has roughly the same
-performance with the random stategy according the average stesp.
+performance with the random strategy according the average steps.
 
-Figure 1 provides a detailed comparision between the three models based on the residual 
+Figure 1 provides a detailed comparison between the three models based on the residual 
 steps of these 50 test instances. In the graph noted in the form of "A vs. B", we obtain
 the residual steps by computing $steps(B) - steps(A)$ for each instances.
-For example, Figure 1 (a) shows the comparision between random strategy and RL with linear 
+For example, Figure 1 (a) shows the comparison between random strategy and RL with linear 
 embedding, the negative numbers on y-axis mean that on these instances the RL model performs
@@
 If we compare the random baseline with graph-based RL, again, we do not observe
-significant improvement. We conjuecture this happens from two reasons: 
+significant improvement. We conjecture this happens from two reasons: 
 1) the graph model although captures the structural information, but it is more 
@@
 may improve the performance;
-2) If the model is trained with more data, we can expect a performance imporvement.
+2) If the model is trained with more data, we can expect a performance improvement.
 
@@
 
-As artifical neural networks greatly succeeded in vision and natural language tasks in last ten years,
+As artificial neural networks greatly succeeded in vision and natural language tasks in last ten years,
 many works that combine deep learning and logical reasoning also has been emerged recently.
@@
 We adapt a similar graph representation from \citeauthor{selsam2018learning}, but use it 
-on DPLL, which poses a difficulty on handling dynamic graph strcuture.
+on DPLL, which poses a difficulty on handling dynamic graph structure.
 Based on \citeauthor{selsam2018learning}'s NeuraSAT, a ICLR 2019 paper
@@
 on larger instances as future works. The graph currently experimented is still 
-static, implementing and tweaking the dynamicaly updated GNN (as discussed in 
+static, implementing and tweaking the dynamically updated GNN (as discussed in 
 Section 2.2) is another on-going work.
